---
title: "Average and threshold motion for 22q and PNC subjects"
author: "Marc Jaskir"
date: "1/22/2023"
output: 
  html_document: 
    toc: yes
---

## Load demographics and list files containing motion data
```{r}
rm(list=ls())

setwd('/Users/mjaskir/ngg/rotations/alexanderbloch/22q/data/motion/22q/')
file_list_22q = list.files(pattern="*.tsv")

setwd('/Users/mjaskir/ngg/rotations/alexanderbloch/22q/data/motion/pnc/')
file_list_pnc = list.files(pattern="*.tsv")

# Load in data frame without motion (to reference subject lists)
demo_22q_no_motion <- read.csv('/Users/mjaskir/ngg/rotations/alexanderbloch/22q/data/demographics/22q/demo_22q_NO_MOTION.csv')
demo_pnc_no_motion <- read.csv('/Users/mjaskir/ngg/rotations/alexanderbloch/22q/data/demographics/pnc/demo_pnc_NO_MOTION.csv')
```

## Exclude data for unused or incomplete 22q subjects/sessions
```{r}
# Cross-reference motion files with 22q demographics dataset
file_list_22q_clean <- c()
for (file in file_list_22q) {
  sub <- strsplit(file,'_')[[1]][1]
  if (sub %in% demo_22q_no_motion$sub) {
    file_list_22q_clean <- c(file_list_22q_clean, file)
  }
}
```

Check which 22q subjects have an unusual number (!= 2) of functional scans
```{r}
skip_indices <- c()
for (sub_22q in demo_22q_no_motion$sub) {
  func_scans <- file_list_22q_clean[grep(sub_22q, file_list_22q_clean)]
  if (length(func_scans) != 2) {
    skip_indices <- c(skip_indices, which(file_list_22q_clean %in% func_scans))
    print(func_scans)
  }
}
```

Excluding sub-016095 ses-008971 (this extra session was previously identified as missing resting state)
```{r}
file_list_22q_clean <- file_list_22q_clean[-skip_indices[3]]
```

## Exclude data for unused or incomplete PNC subjects/sessions
```{r}
file_list_pnc_clean <- c()
for (file in file_list_pnc) {
  sub <- strsplit(file,'_')[[1]][1]
  if (sub %in% demo_pnc_no_motion$sub) {
    file_list_pnc_clean <- c(file_list_pnc_clean, file)
  }
}
```

Check which PNC subjects have an unusual number (!= 2) of functional scans
```{r}
skip_indices <- c()
for (sub_pnc in demo_pnc_no_motion$sub) {
  func_scans <- file_list_pnc_clean[grep(sub_pnc, file_list_pnc_clean)]
  if (length(func_scans) != 2) {
    skip_indices <- c(skip_indices, which(file_list_pnc_clean %in% func_scans))
    print(func_scans)
  }
}
```

Seems like every subject has 2 functional scans - here's another sanity check
```{r}
pnc_fnames_abbrev <- c()
for (file in file_list_pnc_clean) {
  fname_parsed <- strsplit(file,'_')
  if (fname_parsed[[1]][3] == 'task-idemo') {
    pnc_fnames_abbrev <- c(pnc_fnames_abbrev, paste0(fname_parsed[[1]][3], '_', fname_parsed[[1]][4], '_', fname_parsed[[1]][5], '_', fname_parsed[[1]][6], '_', fname_parsed[[1]][7])) 
  } else if (fname_parsed[[1]][3] == 'task-rest') {
    pnc_fnames_abbrev <- c(pnc_fnames_abbrev, paste0(fname_parsed[[1]][3], '_', fname_parsed[[1]][4], '_', fname_parsed[[1]][5], '_', fname_parsed[[1]][6], '_', fname_parsed[[1]][7], '_', fname_parsed[[1]][8])) 
  }
}
table(pnc_fnames_abbrev)
```

- Looks good!

## Load motion data
```{r}
setwd('/Users/mjaskir/ngg/rotations/alexanderbloch/22q/data/motion/22q/')
files_22q = lapply(file_list_22q_clean, read.delim, header=FALSE)
names(files_22q) <- file_list_22q_clean

setwd('/Users/mjaskir/ngg/rotations/alexanderbloch/22q/data/motion/pnc/')
files_pnc = lapply(file_list_pnc_clean, read.delim, header=FALSE)
names(files_pnc) <- file_list_pnc_clean
```

## Compute mean FD for each scan
```{r}
meanFD_summary_22q <- matrix(nrow=length(files_22q),ncol=4)
for (files in names(files_22q)) {
  sub <- strsplit(files,'_')[[1]][1]
  ses <- strsplit(files,'_')[[1]][2]
  scan <- strsplit(files,'_')[[1]][3]
  meanFD <- mean(as.numeric(unlist(files_22q[files])))
  meanFD_summary_22q[which(names(files_22q) == files),] <- c(sub,ses,scan,meanFD)
}
meanFD_summary_22q <- as.data.frame(meanFD_summary_22q)
colnames(meanFD_summary_22q) <- c('sub','ses','scan','meanFD')
meanFD_summary_22q$meanFD <- as.numeric(meanFD_summary_22q$meanFD)

meanFD_summary_pnc <- matrix(nrow=length(files_pnc),ncol=4)
for (files in names(files_pnc)) {
  sub <- strsplit(files,'_')[[1]][1]
  ses <- strsplit(files,'_')[[1]][2]
  scan <- strsplit(files,'_')[[1]][3]
  meanFD <- mean(as.numeric(unlist(files_pnc[files])))
  meanFD_summary_pnc[which(names(files_pnc) == files),] <- c(sub,ses,scan,meanFD)
}
meanFD_summary_pnc <- as.data.frame(meanFD_summary_pnc)
colnames(meanFD_summary_pnc) <- c('sub','ses','scan','meanFD')
meanFD_summary_pnc$meanFD <- as.numeric(meanFD_summary_pnc$meanFD)
```

## Exclude high-motion subjects
Only include subjects for which both scans have mean FD < 0.7, since Cornblath et al 2021 (same 22q sample) set motion threshold at 0.7
```{r}
high_motion_subs_22q <- c()
for (scan in 1:nrow(meanFD_summary_22q)) {
  scan_data <- meanFD_summary_22q[scan,]
  if (scan_data$meanFD >= 0.7) {
    high_motion_subs_22q <- c(high_motion_subs_22q, scan_data$sub)
  }
}

meanFD_summary_22q <- meanFD_summary_22q[! meanFD_summary_22q$sub %in% high_motion_subs_22q,]

high_motion_subs_pnc <- c()
for (scan in 1:nrow(meanFD_summary_pnc)) {
  scan_data <- meanFD_summary_pnc[scan,]
  if (scan_data$meanFD >= 0.7) {
    high_motion_subs_pnc <- c(high_motion_subs_pnc, scan_data$sub)
  }
}

meanFD_summary_pnc <- meanFD_summary_pnc[! meanFD_summary_pnc$sub %in% high_motion_subs_pnc,]
```

## Average motion data within subjects
```{r}
# 22q
meanFD_summary_22q_sublevel <- matrix(nrow=length(unique(meanFD_summary_22q$sub)),ncol=2)
row <- 1
for (subs in unique(meanFD_summary_22q$sub)) {
  sub_data <- meanFD_summary_22q[meanFD_summary_22q$sub==subs,]
  sub_meanFD <- mean(sub_data$meanFD)
  meanFD_summary_22q_sublevel[row,] <- c(subs,sub_meanFD)
  row <- row + 1
}
meanFD_summary_22q_sublevel <- as.data.frame(meanFD_summary_22q_sublevel)
colnames(meanFD_summary_22q_sublevel) <- c('sub','meanFD')

# pnc
meanFD_summary_pnc_sublevel <- matrix(nrow=length(unique(meanFD_summary_pnc$sub)),ncol=2)
row <- 1
for (subs in unique(meanFD_summary_pnc$sub)) {
  sub_data <- meanFD_summary_pnc[meanFD_summary_pnc$sub==subs,]
  sub_meanFD <- mean(sub_data$meanFD)
  meanFD_summary_pnc_sublevel[row,] <- c(subs,sub_meanFD)
  row <- row + 1
}
meanFD_summary_pnc_sublevel <- as.data.frame(meanFD_summary_pnc_sublevel)
colnames(meanFD_summary_pnc_sublevel) <- c('sub','meanFD')
```

## Merge motion data into demographic datasets
```{r}
demo_22q <- merge(demo_22q_no_motion,meanFD_summary_22q_sublevel,by.x = 'sub', by.y='sub', all.y=TRUE)
demo_pnc <- merge(demo_pnc_no_motion,meanFD_summary_pnc_sublevel,by.x = 'sub', by.y='sub', all.y=TRUE)
```


## Save updated demographics and subject lists
```{r}
# Save demographics file
write.table(demo_22q, '/Users/mjaskir/ngg/rotations/alexanderbloch/22q/data/demographics/22q/demo_22q.csv',sep=',',row.names = FALSE)
write.table(demo_pnc, '/Users/mjaskir/ngg/rotations/alexanderbloch/22q/data/demographics/pnc/demo_pnc.csv',sep=',',row.names = FALSE)

# Create output directories
dir.create("/Users/mjaskir/ngg/rotations/alexanderbloch/22q/data/subject_lists/22q", showWarnings = FALSE, recursive = TRUE)
dir.create("/Users/mjaskir/ngg/rotations/alexanderbloch/22q/data/subject_lists/pnc", showWarnings = FALSE, recursive = TRUE)

# Save final subject lists
write.table(demo_22q$sub, '/Users/mjaskir/ngg/rotations/alexanderbloch/22q/data/subject_lists/22q/subject_list_22q.csv',sep=',',row.names = FALSE, col.names = FALSE)
write.table(demo_pnc$sub, '/Users/mjaskir/ngg/rotations/alexanderbloch/22q/data/subject_lists/pnc/subject_list_pnc.csv',sep=',',row.names = FALSE, col.names = FALSE)
```

